---
title: "IMMORTOOL-use-from-R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IMMORTOOL-use-from-R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# General background

The README of the GitHub repo at https://github.com/petedodd/IMMORTOOL describes how to run the Shiny app version of this package, as well as how to access a live snapshot of this app online without installation.

This vignette will describe how to use the IMMORTOOL package from the commandline or from scripts in R.


First install the library (see README), and then load at the start of a script or R session:

```{r setup}
library(IMMORTOOL)
```


For the example below we will set a random number seed, and take the cohort size to be 0.5 million. Restults should be checked to verify random noise is not affecting them at the desired level of accuracy.
```{r}
cohort.size <- 5e6 #stable to 2 dp
set.seed(1234)
```

The examples below are taken from the article describing this tool. The first set of articles are studies where both naive and more sophisticated analyses have been undertaken. The second set of articles (with various sensitivity analyses around data interpretation) have only naive analyses, and underly the forst plot in our main article.


# First set of examples


## van der Vaart

```{r,fig.dim=c(7,6)}
## Main result (demonstration naive analysis)
ans.vanderVaart <- makeResultList(
                      mortality.times = c(7,30,90),
                      mortality.fracs = c(0.05,0.21,0.31),
                      treatment.times = c(6,9,13,90),
                      treatment.fracs = c(0.09,0.19,0.28,0.37),
                      N = 147, Tmax = 90, Tlandmark = 1,
                      plotfit = TRUE,
                      simulation.cohort.size = cohort.size)

```

## Jones and Fowler

```{r,fig.dim=c(7,6)}

## mortality
tzm <- c(1,2,3,4,5,6,7,8,9,10,11,12,90)
fzm <- c(0.01,0.04,0.05,0.05,0.06,0.07,0.08,0.08,0.09,0.11,0.12,0.13,0.20)

## treatment
tzt <- c(1,2,45)
fzt <- c(0.73,0.84,0.93)

## Main result 
ans.JonesandFowler <- makeResultList(mortality.times = tzm,
                                     mortality.fracs = fzm,
                                     treatment.times = tzt,
                                     treatment.fracs = fzt,
                                     N = 117, Tmax = 90, Tlandmark = 1,
                                     plotfit = TRUE,
                                     simulation.cohort.size = cohort.size)

```
NOTE should we have Tearly=2? But these analyses not used?


## Geleris

```{r,fig.dim=c(7,6)}
tzm <- c(1,30)
fzm <- c(0.05,0.25)
tzt <- c(0.1,1,9)
fzt <- c(0.27,0.51,0.59)

## Standard assumptions
ans.Geleris <- makeResultList(mortality.times = tzm,
                              mortality.fracs = fzm,
                              treatment.times = tzt,
                              treatment.fracs = fzt,
                              N=346,Tmax=30,Tlandmark=2,
                              plotfit = TRUE,
                              simulation.cohort.size = cohort.size)
```


## Combined

```{r}
## Calculate confidence intervals: make list of all results
nmz <- c('van der Vaart','Jones and Fowler','Geleris')
L <- list(ans.vanderVaart,ans.JonesandFowler,ans.Geleris)
names(L) <- nmz

## create table from results list
tab <- makeCItable(L)

## show just the id and RR+CIs:
knitr::kable(tab[,c('id','CI.a','CI.b','CI.c','CI.d')])

```

NOTE could modify above to grey out the irrelevant cells?


# Second set of examples


```{r}
Tmax <- 30
Tlandmark <- 1
```


## Adalat

```{r,fig.dim=c(7,6)}
## Standard assumptions
ans.Adalat.standard <- makeResultList(mortality.times = c(2,8,30),
                                      mortality.fracs = c(0.07,0.14,0.14),
                                      treatment.times = c(0.5,30),
                                      treatment.fracs = c(0.19,0.38),
                                      N = 3, Tmax = Tmax, Tlandmark = Tlandmark,
                                      plotfit = TRUE,
                                      simulation.cohort.size = cohort.size)

## Low ITB assumptions
ans.Adalat.low <- makeResultList(mortality.times = c(2,8,30),
                                 mortality.fracs = c(0.05,0.14,0.14),
                                 treatment.times = c(0.25,30),
                                 treatment.fracs = c(0.19, 0.38),
                                 N = 3, Tmax = Tmax, Tlandmark = Tlandmark,
                                 simulation.cohort.size = cohort.size)

## High ITB assumptions
ans.Adalat.high <- makeResultList(mortality.times = c(2,8,30),
                                  mortality.fracs = c(0.10,0.14,0.14),
                                  treatment.times = c(1, 30),
                                  treatment.fracs = c(0.19, 0.38),
                                  N = 3, Tmax = Tmax, Tlandmark = Tlandmark,
                                  simulation.cohort.size = cohort.size)


```


## Carapetis

```{r,fig.dim=c(7,6)}

## Standard assumptions
ans.Carapetis.standard <- makeResultList(mortality.times = c(2,8,30),
                                         mortality.fracs = c(0.09,0.19,0.19),
                                         treatment.times = c(0.5,30),
                                         treatment.fracs = c(0.18,0.35),
                                         N = 7, Tmax = Tmax, Tlandmark = Tlandmark,
                                         plotfit = TRUE,
                                         simulation.cohort.size = cohort.size)

## Low ITB assumptions
ans.Carapetis.low <- makeResultList(mortality.times = c(2,8,30),
                                    mortality.fracs = c(0.06,0.19,0.19),
                                    treatment.times = c(0.25,30),
                                    treatment.fracs = c(0.18,0.35),
                                    N = 7, Tmax = Tmax, Tlandmark = Tlandmark,
                                    simulation.cohort.size = cohort.size)

## High ITB assumptions
ans.Carapetis.high <- makeResultList(mortality.times = c(2,8,30),
                                     mortality.fracs = c(0.13,0.19,0.19),
                                     treatment.times = c(1,30),
                                     treatment.fracs = c(0.18,0.35),
                                     N = 7, Tmax = Tmax, Tlandmark = Tlandmark,
                                     simulation.cohort.size = cohort.size)


```



## Linner

```{r,fig.dim=c(7,6)}

## Standard assumptions
ans.Linner.standard <- makeResultList(mortality.times = c(2,8,30),
                                      mortality.fracs = c(0.13,0.27,0.27),
                                      treatment.times = c(0.5,30),
                                      treatment.fracs = c(0.2,0.4),
                                      N = 14, Tmax = Tmax, Tlandmark = Tlandmark,
                                      plotfit = TRUE,
                                      simulation.cohort.size = cohort.size)

## Low ITB assumptions
ans.Linner.low <- makeResultList(mortality.times = c(2,8,30),
                                 mortality.fracs = c(0.09,0.27,0.27),
                                 treatment.times = c(0.25,30),
                                 treatment.fracs = c(0.2,0.4),
                                 N = 14, Tmax = Tmax, Tlandmark = Tlandmark,
                                 simulation.cohort.size = cohort.size)

## High ITB assumptions
ans.Linner.high <- makeResultList(mortality.times = c(2,8,30),
                                  mortality.fracs = c(0.18,0.27,0.27),
                                  treatment.times = c(1,30),
                                  treatment.fracs = c(0.2,0.4),
                                  N = 14, Tmax = Tmax, Tlandmark = Tlandmark,
                                  simulation.cohort.size = cohort.size)

```



## Combined

```{r}
### Calculate confidence intervals: make list of all results
nmz <- c('Adalat.standard','Carapetis.standard','Linner.standard',
         'Adalat.low','Carapetis.low','Linner.low',
         'Adalat.high','Carapetis.high','Linner.high')
L <- list(ans.Adalat.standard,ans.Carapetis.standard,ans.Linner.standard,
          ans.Adalat.low,ans.Carapetis.low,ans.Linner.low,
          ans.Adalat.high,ans.Carapetis.high,ans.Linner.high)
names(L) <- nmz


## create table from results list
tab <- makeCItable(L)

## show just the id and RR+CIs for naive only:
knitr::kable(tab[,c('id','CI.a')])

```

# Examples using internals of above

The above `makeResultList` function is a wrapper to simplify workflows, and for use in the Shiny app. It may be useful and instructuve to understand the steps that this function is carrying out.

Suppose we consider the data from Andalat et al.:
```{r}
mortality.times <- c(2,8,30)
mortality.fracs <- c(0.07,0.14,0.14)
treatment.times <- c(0.5,30)
treatment.fracs <- c(0.19,0.38)
```

We can use these data to generate a fit to mortality parameters, and then fit the treatment/exposure hazard given the mortality parameters, before combining the 4 Weibull parameters into a single list for convenience:
```{r}
## mortality fit
mortality.parms <- Yfit(mortality.times, mortality.fracs)

## treatment data and fit
treatment.data <- cbind(treatment.times, treatment.fracs)
treatment.parms <- getTxParz(treatment.data, mortality.parms)

## combine
W <- list(k.d=mortality.parms[1], L.d=mortality.parms[2],
              k.e=treatment.parms$k.e, L.e=treatment.parms$L.e)

```

These parameters can then be fed to the main function than creates and runs cohorts, before returning a list of results:
```{r}

outputs <- ITBstats(
    N = cohort.size,
    Tstop = Tmax, Tlandmark = 2, Tearly = 1,
    rtt.exposure = function(n) rweibull(n, W$k.e, W$L.e),
    rtt.death = function(n) rweibull(n, W$k.d, W$L.d),
    rtt.ltfu = function(n) rweibull(n, 1, 36500)
  )

```

Note: the function `ITBstats` can take any random number generator distributions as inputs, so does not need to restrict to Weibull distributions.